|

### 2020-05-16 paper/Manuscript.Rmd
edits and suggestion by Anna-Lena

---
title: |
    | WORCS: A Workflow for Open Reproducible Code in Science
    | Version 0.1.0 (Feb 8, 2020)

@@

author:
  - name: "Caspar J. Van Lissa"
    institution: "1,2"
    corresponding: yes
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
  - name: "Andreas M. Brandmaier"
    institution: "3,4"
  - name: "Loek Brinkman"
    institution: "2,5"
  - name: "Aaron Peikert"
    institution: "3,6"
  - name: "Barbara Vreede"
    institution: "7"

affiliation:
  - id: "1"
    institution: "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id: "2"
    institution: "Open Science Community Utrecht"
  - id: "3"
    institution: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id: "4"
    institution: "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
  - id: "5"
    institution: "UMC Utrecht"
  - id: "6"
    institution: "Humboldt-Universitaet zu Berlin"
  - id: "7"
    institution: "University Library Utrecht"

@@

@@

Benefits to individual scientists: Automation (after a learning curve), thus more productive, avoid mistakes, join projects, clone projects and make them your own.
Collective benefits: Cumulative science, -->

@@

The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by journals, funders, and organizations. However, large organizations are slow to change. For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
<!-- A more recent preprint reported that many journals do not explicitly state their policies about  klebelPeerReviewPreprint2020 --> At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012]. Many individual researchers are willing to adopt best practices for open science, even in the absence of institutional support. The formation of "Open Science Communities" at universities around the globe illustrates this grassroots movement. This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of open science. This workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them.

@@

@@

WORCS relies on several free, open source software solutions. The first of these is *dynamic document generation*. Dynamic document generation (DDG) means writing scientific reports in a format that interleaves written reports with blocks of computer code used to conduct the analyses. The text is automatically formatted as an APA-style manuscript [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020], and results of the code blocks are automatically generated and insterted in the text, or rendered as Figures and Tables. Dynamic document generation supercedes the classical approach of using separate programs to write prose and conduct analyses, and then manually copy-pasting analysis results into the text. Although there is a slight learning curve to transitioning to DDG, the investment really pays off:

* The time saved from painstakingly copy-pasting output and manually formatting text soon outweighs the investment of switching to a new program
* Human error in copy-pasting results is eliminated
* When revisions require major changes to the analyses, all results, Figures and Tables are automatically updated
* Flexibility in output formats means that a manuscript can be rendered to presentation format, or even to a website. 
* Reproducibility is guaranteed, and can be verified by reviewers and/or readers, because the code is run each time the document is compiled.

@@

The second solution is *version control*: Maintaining an indelible log of every change to all project files. Version control is a near-essential tool for scientific reproducibility, as anyone learns who has had the experience of accidentally deleting a crucial file, or of being unable to reproduce analyses because they ran some analyses interactively instead of using syntax [see also @blischakQuickIntroductionVersion2016]. Many scientists use some form of *implicit* version control; for example, by renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"), tracking changes in word processing software, or using cloud hosting services that retain backups of previous versions. WORCS instead uses the *explicit* version control software "Git". Git tracks changes to files, and stores these changes when the user makes a "commit". Git retains a complete log of all commits, and users can compare changes between different commits, or go back to a previous version of the code (for example, after making a mistake, or to replicate a previous version of the results). A project version controlled with Git is called a "Repository", or Repo. It is also possible to prevent some files from being version controlled - such as privacy sensitive data. The `.gitignore` file is a list of files that should not be version-controlled, and thus exist only on the user's private computer.

@@

<!-- LB: and onwards: this section requires some background knowledge on R/packages/etc. This could be a place where readers stumble and stop reading. Perhaps introduce the concept of the R project template earlier, to see section on dependencies more into perspective. 
[main issue] it should become clear earlier what the WORCS workflow entails: e.g. switching to working with a particular R Project Template. Discuss the ‘workflow' section earlier. Related: Can you make a schematic overview of the WORCS?overview--> 
Many solutions exist to ensure computational reproducibility. These solutions typically work by enveloping your research project in a distinct "environment" that only has access to programs that are explicitly installed, and maintaining a record of these programs. These solutions differ in user-friendliness and effectiveness. If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine. This is effective, and also pertains to software outside the R environment (e.g., preprocessing tools for neuroimaging data). However, it is very difficult to set up. <!--For small projects, or novice users || Mention CodeOcean here-->A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project. This solution is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages must be installed anew, which can take a long time (sometimes hours) and a lot of hard drive space. Moreover, Git/GitHub is not designed to store your entire project library, so `packrat` does not address the issue of how other users are supposed to get access to your library. In conclusion, `packrat` is somewhat unwieldy. The package `checkpoint` offers an extremely lightweight solution: Instead of storing all software used, it simply records the calendar date, and then installs the version of all required packages that was available on the central R repository `CRAN` on that date from an archival website. Thus, all that is needed to restore a project is the date on which it was conducted. This is very user-friendly, and reasonably effective - the only problem is that this solution is limited to `CRAN` packages, and cannot mix package versions from different dates.

@@

A few additional considerations were considered in developing WORCS. The first consideration is to develop a workflow for R, instead of for SPSS or Python. Several arguments support the choice to work in R. First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for open science. Second, all of the aforementioned tools for an open science workflow are already implemented in R. Many statistical methods are first implemented in R, as there is a vibrant online community developing new R-packages and providing support and tutorials for existing ones. Wrappers in R are often available for tools developed in other programming languages. <!-- LB: Perhaps give some pointers to learn R (e.g. R4DataScience)-->Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper. Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.

@@

Another key consideration is what filetypes to use. A case can be made to use text-based files only (or as much as possible), instead of binary files. Text-based files can be read by machines and humans alike, without requiring special (licensed) software. Binary files, such as Word (`.docx`) or SPSS (`.sav`, `.spo`),  must be decoded by software first. Git is designed to keep track of changes to the content of text-based files. Each change reflects a word or line of code added or removed. Binary files, on the other hand, should not be version controlled. The changes in the encrypted file are uninterpretable. Moreover, text-based files are often smaller than binary files, which means they take up less space on cloud hosting services, such as GitHub. Uploading large binary files is frowned upon. GitHub also allows visitors to peruse text-based files online, and conveniently renders certain filetypes. For example, GitHub highlights syntax in `.R` and `.Rmd` files, and displays `.csv` files as spreadsheets.

@@

<!-- LB:  this section comes at a weird place in the flow of the paper. -->Sometimes, open science principles are discussed as a kind of "antidote" to scientific misconduct. Yet these principles also present an opportunity to conduct more reliable research, with fewer mistakes, more convenient collaboration, more effective internal and external collaboration, and the potential for crowd-sourced error correction.

@@

When using R, software (R packages) can be conveniently cited by calling `citation("packagename")`. This returns an APA-style reference, and a BibTeX entry. However, users of the WORCS project template for RStudio do not need to manually cite R-packages. WORCS generates a template manuscript in APA6 style, based on the `papaja` package [@austPapajaPrepareReproducible2020], which automatically cites all packages used.

@@

It is easiest to mark the distinction between essential and non-essential references from the start, instead of going back to cut references prior to publication. Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@`), used to cite a non-essential reference. Users can render a manuscript to PDF either with, or without, comprehensive citations.

@@

@@

# The workflow

@@

Another important challenge is managing collaborations when only the lead author uses Rmarkdown, and the coauthors use Word. In this case, it is possible to Knit the manuscript to Word (.docx), by changing the line `output: papaja::apa6_pdf` in the manuscript's YAML header to `output: papaja::apa6_docx`. There are some limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents). When soliting feedback from co-authors, ask them to use Tracked Changes and comment bubbles in Word. <!-- LB: The work-around with track-changes in Word files is doable, but not very satisfying. I think you can also use Overleaf instead of Word (user friendly and web based) for bidirectional communication, but haven’t tried this myself. I didn’t know about ‘redoc’, will check it out.-->Then, manually incorporate their changes into the `manuscript.Rmd` file. In most cases, this is the most user-friendly approach, as most lead authors would review changes by co-authors anyway. A second approach is to ask collaborators to work in plain text. In this case, send collaborators the `manuscript.Rmd` file, and ask them to open (and save) it in Word or Notepad as a plain text file. When they send it back, make sure any changes to your local file are committed, and then simply overwrite your local version with their file. In RStudio, select the file in the Git tab, and click the Diff button to examine what changes the collaborator has made relative to your last committed version. Finally, @peikertReproducibleDataAnalysis2019 mention that the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] might be useful for collaborating with Word users. This package enables bidirectional conversion between Word and Rmarkdown, with support for tracked changes. However, a simpler solution (i.e., asking collaborators to work in plain text) might be sufficient.

If all collaborators are committed to using WORCS, they can [Fork the repository](https://help.github.com/en/github/getting-started-with-github/fork-a-repo) from the lead author on GitHub, [clone it to their local device](https://help.github.com/en/github/creating-cloning-and-archiving-repositories/cloning-a-repository), make their own changes, and [send you a pull request](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) to incorporate their changes. Working this way is extremely conducive to scientific collaboration [@ramGitCanFacilitate2013]. Moreover, being familiar with GitHub opens doors to new forms of collaboration: In the open source software community, continuous peer review and voluntary collaborative acts by strangers who are interested in a project are commonplace [see @adolphOpenBehavioralScience2012].  This kind of collaboration is budding in scientific software development as well; for example, I became a co-author on several R-packages after submitting pull requests with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018]. It is also possible to invite such collaboration by [opening Issues](https://help.github.com/en/github/managing-your-work-on-github/about-issues) for tasks that still need to be accomplished, and tag known Collaborators to address them, or invite external collaborators to contribute their expertise.

>>>>

---
title: |
    | WORCS: A Workflow for Open Reproducible Code in the Social Sciences and Beyond <!-- AL: I strongly recommend to make the title more specific. "Science" is very general, and in the Anglo-American understanding actually refers to the beta sciences. Also, the abstract and the paper start with talking about the situation in the social sciences and derive the motivaion from there, so it would be good if the title would reflect this. The acronym can stay as it is. ;) -->
    | Version 0.1.0 (Feb 8, 2020)

@@

author:
  - name: "Caspar J. Van Lissa"
    institution: "1,2"
    corresponding: yes
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
  - name: "Andreas M. Brandmaier"
    institution: "3,4"
  - name: "Loek Brinkman" 
    institution: "2,5"
  - name: "Aaron Peikert"
    institution: "3,6"
  - name: "Barbara Vreede"
    institution: "7"
  - name: "Anna-Lena Lamprecht"
    institution: "2,8"

affiliation:
  - id: "1"
    institution: "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id: "2"
    institution: "Open Science Community Utrecht"
  - id: "3"
    institution: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id: "4"
    institution: "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
  - id: "5"
    institution: "UMC Utrecht"
  - id: "6"
    institution: "Humboldt-Universitaet zu Berlin"
  - id: "7"
    institution: "University Library Utrecht"
  - id: "8"
    institution: "Utrecht University, Department of Information and Computing Sciences"

@@

<!-- AL: I second LB's comment somewhere below that the overall WORCS workflow should be introduced much earlier, maybe already here. A diagrammatic overview/flowchart of the principal steps would be great. It should catch the reader's eye and give an immediate idea of what WORCS is about. Now it is indeed quite buried in the details that follow. -->

@@

Benefits to individual scientists: Automation (after a learning curve), thus more productive, avoid mistakes, join projects, clone projects and make them your own.
Collective benefits: Cumulative science, -->
<!-- AL: These "value propositions" would also be good to discuss already in the introduction. Directly after the introduction of the WORCS overview would make sense, I think. -->

The remainder of the paper is structured as follows. ... <!-- AL: Or some variation of this obligatory end-of-intro paragraph. As the paper touches on a lot of aspects, from principal to very practical, it would be very useful to provide some guidance here. -->

@@

The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by journals, funders, and organizations. However, large organizations are slow to change. For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
<!-- A more recent preprint reported that many journals do not explicitly state their policies about  klebelPeerReviewPreprint2020 --> At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012]. Many individual researchers are willing to adopt best practices for open science, even in the absence of institutional support. The formation of "Open Science Communities" at universities around the globe illustrates this grassroots movement. <!-- AL: Is there a reference for this? Maybe a suitable blog post, or just the OSC website? --> This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of open science. The WORCS workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them.

@@

<!-- AL: Similar to my comment on the title, in this section it needs to make clear which scope of related work it covers. For example, in the life sciences there has also been a lot of work on practices for increasing reproducibility, with automated workflows, containerization, Research Objects, provenance protocols, and a lot more. I don't think it's necessary for this paper to discuss this, but that is under the provision that it defines its scope better. -->

@@

WORCS relies on several free, open source software solutions. The first of these is *dynamic document generation*. <!-- AL: A bit nitpicky, but: DDG as such is a technique, not a solution. Could also be implemented as proprietary, closed source. Papaja is the tool/solution. The description of formatting as an APA-style manuscript also seems to be tool-specific (generall DDG could also generate any other kind of format), so maybe reorganize this paragraph. --> Dynamic document generation (DDG) means writing scientific reports in a format that interleaves written reports with blocks of computer code used to conduct the analyses. The text is automatically formatted as an APA-style manuscript [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020], and results of the code blocks are automatically generated and insterted in the text, or rendered as Figures and Tables. Dynamic document generation supercedes the classical approach of using separate programs to write prose and conduct analyses, and then manually copy-pasting analysis results into the text. Although there is a slight learning curve to transitioning to DDG, the investment really pays off:

* The time saved from painstakingly copy-pasting output and manually formatting text soon outweighs the investment of switching to a new program
* Human error in copy-pasting results is eliminated
* When revisions require major changes to the analyses, all results, figures and tables are automatically updated
* Flexibility in output formats means that a manuscript can be rendered to presentation format, or even to a website. 
* Reproducibility is guaranteed, and can be verified by reviewers and/or readers, because the code is run each time the document is compiled.

@@

The second solution is *version control*: Maintaining an indelible log of every change to all project files. Version control is a near-essential tool for scientific reproducibility, as anyone learns who has had the experience of accidentally deleting a crucial file, or of being unable to reproduce analyses because they ran some analyses interactively instead of using syntax <!-- AL: syntax is also to be respected in e.g. interactive R environments, so that alone does not help -  maybe "code files" instead? -->  [see also @blischakQuickIntroductionVersion2016]. Many scientists use some form of *implicit* version control; for example, by renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"), tracking changes in word processing software, or using cloud hosting services that retain backups of previous versions. WORCS instead uses the *explicit* version control software "Git". Git tracks changes to files, and stores these changes when the user makes a "commit". Git retains a complete log of all commits, and users can compare changes between different commits, or go back to a previous version of the code (for example, after making a mistake, or to replicate a previous version of the results). A project version controlled with Git is called a "Repository", or Repo. It is also possible to prevent some files from being version controlled - such as privacy sensitive data. The `.gitignore` file is a list of files that should not be version-controlled, and thus exist only on the user's private computer.

@@

<!-- LB: and onwards: this section requires some background knowledge on R/packages/etc. This could be a place where readers stumble and stop reading. Perhaps introduce the concept of the R project template earlier, to see section on dependencies more into perspective. 
[main issue] it should become clear earlier what the WORCS workflow entails: e.g. switching to working with a particular R Project Template. Discuss the ‘workflow' section earlier. Related: Can you make a schematic overview of the WORCS?overview--> 
Many solutions exist to ensure computational reproducibility. These solutions typically work by enveloping your research project in a distinct "environment" that only has access to programs that are explicitly installed, and maintaining a record of these programs. These solutions differ in user-friendliness and effectiveness. If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a container <!-- AL: Docker containers are not virtual machine. Actually that is the whole idea of Docker (and other container technologies), as real virtual machines need way more resources. --> . This is effective, and also pertains to software outside the R environment (e.g., preprocessing tools for neuroimaging data). However, it is very difficult to set up. <!--For small projects, or novice users || Mention CodeOcean here-->A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project. This solution is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages must be installed anew, which can take a long time (sometimes hours) and a lot of hard drive space. Moreover, Git/GitHub is not designed to store your entire project library, so `packrat` does not address the issue of how other users are supposed to get access to your library. In conclusion, `packrat` is somewhat unwieldy. The package `checkpoint` offers an extremely lightweight solution: Instead of storing all software used, it simply records the calendar date, and then installs the version of all required packages that was available on the central R repository `CRAN` on that date from an archival website. Thus, all that is needed to restore a project is the date on which it was conducted. This is very user-friendly, and reasonably effective - the only problem is that this solution is limited to `CRAN` packages, and cannot mix package versions from different dates.

@@

A few additional considerations were considered in developing WORCS. The first consideration is to develop a workflow for R, instead of for SPSS or Python. Several arguments support the choice to work in R. First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for open science <!-- AL: Python is also open source, so phrase this argument with care. --> . Second, all of the aforementioned tools for an open science workflow are already implemented in R. Many statistical methods are first implemented in R, as there is a vibrant online community developing new R-packages and providing support and tutorials for existing ones. Wrappers in R are often available for tools developed in other programming languages. <!-- AL: Maybe this should be the first argument. While Python is catching up, R still is THE statistical programming environment in many communities, and is thus favorable in many cases.--> <!-- LB: Perhaps give some pointers to learn R (e.g. R4DataScience)-->Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012]<!-- AL: If the reference is really from 2012, it is not a good support for "at the time of writing"... --> , following SPSS, which has no support for any of the open science tools discussed in this paper. Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.<!-- AL: Many of these are also true for Python. There, Jupyter notebooks are very popular as export and exchange format (also using MarkDown), might be good to refer to that somewhere. I think they also support R. -->
<!-- AL: And maybe this goes a bit too far for this paper, but https://www.scientificpaperofthefuture.org/ also fits in this context. -->

@@

Another key consideration is what filetypes to use. A case can be made to use text-based files only (or as much as possible), instead of binary files. Text-based files can be read by machines and humans alike (just requiring a plain text editor), without requiring special (often licensed) software. Binary files, such as Word documents or (`.docx`) or SPSS objects (`.sav`, `.spo`), must be decoded by specialized software first. Git is designed to keep track of changes to the content of text-based files. Each change reflects a word or line of code added or removed. Binary files, on the other hand, should not be version controlled. The changes in the encrypted file are uninterpretable. Moreover, text-based files are often smaller than binary files, which means they take up less space on cloud hosting services, such as GitHub. Uploading large binary files is frowned upon. GitHub also allows visitors to peruse text-based files online, and conveniently renders certain filetypes. For example, GitHub highlights syntax in `.R` and `.Rmd` files, and displays `.csv` files as spreadsheets.

@@

<!-- LB:  this section comes at a weird place in the flow of the paper. --><!-- AL: +1 --> Sometimes, open science principles are discussed as a kind of "antidote" to scientific misconduct. Yet these principles also present an opportunity to conduct more reliable research, with fewer mistakes, more convenient collaboration, more effective internal and external collaboration, and the potential for crowd-sourced error correction.

@@

<!-- AL: Some general things about software citation that could be good to mention here: Tool or software papers are also common, often as short papers, sometimes called "Application Notes". The Journal of Open Source Software (JOSS) is a way of turning software directly into a publication, without the need of writing an actual paper about it (the review is about the code). https://citation-file-format.github.io/ is a file format to include software citation information into source code repositories.--> 
When using R, software (R packages) can be conveniently cited by calling `citation("packagename")`. This returns an APA-style reference, and a BibTeX entry. However, users of the WORCS project template for RStudio do not need to manually cite R-packages. WORCS generates a template manuscript in APA6 style, based on the `papaja` package [@austPapajaPrepareReproducible2020], which automatically cites all packages used.

@@

It is easiest to mark the distinction between essential and non-essential references from the start, instead of going back to cut references prior to publication. Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@`)<!-- AL: what is the difference? do you mean '@@'? or is it just showing the same on my machine?-->, used to cite a non-essential reference. Users can render a manuscript to PDF either with, or without, comprehensive citations.

@@

<!-- Maybe remark that in the "closed" case, a the data should be backupped somewhere else, as GitHub will not do the job in this setting. -->

@@

# The workflow 
<!-- AL: Agreeing with LB's ealier comment, it would be good to discuss this overall workflow before going into the details. -->

@@

Another important challenge is managing collaborations when only the lead author uses Rmarkdown, and the coauthors use Word. In this case, it is possible to Knit the manuscript to Word (.docx), by changing the line `output: papaja::apa6_pdf` in the manuscript's YAML header to `output: papaja::apa6_docx`. There are some limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents). When soliting feedback from co-authors, ask them to use Tracked Changes and comment bubbles in Word. <!-- LB: The work-around with track-changes in Word files is doable, but not very satisfying. I think you can also use Overleaf instead of Word (user friendly and web based) for bidirectional communication, but haven’t tried this myself. I didn’t know about ‘redoc’, will check it out.-->Then, manually<!--this should really be avoided if possible! no fun... --> incorporate their changes into the `manuscript.Rmd` file. In most cases, this is the most user-friendly approach, as most lead authors would review changes by co-authors anyway. A second approach is to ask collaborators to work in plain text. In this case, send collaborators the `manuscript.Rmd` file, and ask them to open (and save) it in Word or Notepad as a plain text file. When they send it back, make sure any changes to your local file are committed, and then simply overwrite your local version with their file. In RStudio, select the file in the Git tab, and click the Diff button to examine what changes the collaborator has made relative to your last committed version. Finally, @peikertReproducibleDataAnalysis2019 mention that the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] might be useful for collaborating with Word users. This package enables bidirectional conversion between Word and Rmarkdown, with support for tracked changes. However, a simpler solution (i.e., asking collaborators to work in plain text) might be sufficient.

If all collaborators are committed to using WORCS, they can [Fork the repository](https://help.github.com/en/github/getting-started-with-github/fork-a-repo) from the lead author on GitHub, [clone it to their local device](https://help.github.com/en/github/creating-cloning-and-archiving-repositories/cloning-a-repository), make their own changes, and [send a pull request](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) to incorporate their changes. Working this way is extremely conducive to scientific collaboration [@ramGitCanFacilitate2013]. Moreover, being familiar with GitHub opens doors to new forms of collaboration: In the open source software community, continuous peer review and voluntary collaborative acts by strangers who are interested in a project are commonplace [see @adolphOpenBehavioralScience2012].  This kind of collaboration is budding in scientific software development as well; for example, I <!-- AL: which of the meanwhile several authors? :) -->  became a co-author on several R-packages after submitting pull requests with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018]. It is also possible to invite such collaboration by [opening Issues](https://help.github.com/en/github/managing-your-work-on-github/about-issues) for tasks that still need to be accomplished, and tag known Collaborators to address them, or invite external collaborators to contribute their expertise.
